---
title: "Bob Ross"
author: "David Moste"
date: "3/7/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
I used the Bob Ross dataset for this tidying project.

### Code
I started by loading the required libraries and the dataset.
```{r}
# load required libraries
library(tidyverse)

# load the bob ross csv
br <- read.csv("bob_ross.csv", header = TRUE)
```

Next, I tidied the data using pivot_longer, filter, and separate.
```{r}
# tidy the bob ross csv (pivot_longer, filter, separate, select, group_by, mutate)
br <- br %>%
  pivot_longer(c("APPLE_FRAME":"WOOD_FRAMED"),
               names_to = "element",
               values_to = "present") %>%
  filter(present == 1) %>%
  separate("EPISODE", into = c("season", "episode"), sep = "E") %>%
  separate("season", into = c("S", "season"), sep = "S") %>%
  select(-S)

br$season <- as.integer(br$season)
br$episode <- as.integer(br$episode)
```

Finally, I created a new dataframe with based on the most common elements in the dataset and graphed this.
```{r}
# create a data frame containing only one instance of each object
most_common <- br %>%
  group_by(element) %>%
  mutate("total" = sum(present)) %>%
  distinct(element, total) %>%
  filter(total > 50)

# plot the most popular elements
ggplot(most_common, aes(x = reorder(element, total), y = total)) +
  geom_bar(stat = "identity") +
  labs(x = "Element",
       y = "Total Appearances") +
  coord_flip()
```

# Conclusion
Based on my analysis,Bob Ross looked at a lot of paintings with trees. Tree and Trees were the two most common words in the dataset.