---
title: "Basic Loading and Transformations"
author: "David Moste"
date: "2/1/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction


## Getting the Data
I started by loading the .csv files into two data frames.
```{r}
cable_data <- data.frame(read.csv("media-mentions-2020/cable_weekly.csv"))
online_data <- data.frame(read.csv("media-mentions-2020/online_weekly.csv"))
```
The next step was to combine the relevant columns into a new data frame titled all_data and provide relevant column names.
```{r}
all_data <- data.frame(cable_data$date, cable_data$name, 
                       cable_data$pct_of_all_candidate_clips,
                       online_data$pct_of_all_candidate_stories)
names(all_data) <- c("Date","Name","Clips","Stories")
```
Finally, I subsetted the data to only include mentions of Joe Biden. At the same time, I converted the Date column into Date objects.
```{r}
all_biden <- subset(all_data, Name == "Joe Biden")
all_biden$Date <- as.Date(all_biden$Date)
```
## Looking at the Data
I started here by looking at the enitre set of data.
```{r}
plot(all_biden$Date, all_biden$Clips + all_biden$Stories,
     type = "l", main = "Biden in the Media", xlab = "Date",
     ylab = "% of Mentions")
```
After looking at the data, it's clear that there is a sharp decrease in Joe Biden mentions starting somewhere in September With this in mind, I subsetted the data again, looking just at mentions from September on.
```{r}
all_biden <- subset(all_biden, Date >= "2019-09-01")
plot(all_biden$Date, all_biden$Clips + all_biden$Stories,
     type = "l", main = "Biden in the Media", xlab = "Date",
     ylab = "% of Mentions")
```

# Conclusions
