---
title: "Sentiment Analysis"
author: "David Moste"
date: "3/31/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
For this assignment, I am tasked with getting an example from *Text Mining with R* running and then extending the example to a new corpus and a neew sentiment lexicon. Sections 1-6 are directly from *Text Mining with R*^1^. Section 7 is my extension.

## 1. Sentiments Dataset
```{r}
library(tidytext)

get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```

## 2. Sentiment Analysis with Inner Join
```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(linenumber = row_number(),
         chapter = cumsum(str_detect(text, regex("^chapter [\\divxlc]", 
                                                 ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)

library(tidyr)

jane_austen_sentiment <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

library(ggplot2)

ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## 3. Comparing the Three Dictionaries
```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice

afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(pride_prejudice %>% 
                            inner_join(get_sentiments("bing")) %>%
                            mutate(method = "Bing et al."),
                          pride_prejudice %>% 
                            inner_join(get_sentiments("nrc") %>% 
                                         filter(sentiment %in% c("positive", 
                                                                 "negative"))) %>%
                            mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)

bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")

get_sentiments("nrc") %>% 
     filter(sentiment %in% c("positive", 
                             "negative")) %>% 
  count(sentiment)

get_sentiments("bing") %>% 
  count(sentiment)
```

## 4. Most Common Positive & Negative Words
```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

bing_word_counts %>%
  group_by(sentiment) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(word, n, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(y = "Contribution to sentiment",
       x = NULL) +
  coord_flip()

custom_stop_words <- bind_rows(tibble(word = c("miss"), 
                                          lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```

## 5. Wordclouds
```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("gray20", "gray80"),
                   max.words = 100)
```

## 6. Looking Beyond Words
```{r}
PandP_sentences <- tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence, text, token = "sentences")

PandP_sentences$sentence[2]

austen_chapters <- austen_books() %>%
  group_by(book) %>%
  unnest_tokens(chapter, text, token = "regex", 
                pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
  ungroup()

austen_chapters %>% 
  group_by(book) %>% 
  summarise(chapters = n())

bingnegative <- get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
  summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
  group_by(book, chapter) %>%
  summarize(negativewords = n()) %>%
  left_join(wordcounts, by = c("book", "chapter")) %>%
  mutate(ratio = negativewords/words) %>%
  filter(chapter != 0) %>%
  top_n(1) %>%
  ungroup()
```

## 7. My Extension
For this section, I analyzed *The Iliad* using three different sentiment lexicons: SentiWordNet^2^, Hu Liu^3^, and SlangSD^4^.
```{r}
# open libraries
library(lexicon)
library(readr)

# open four lexicons: SenticNet, SentiWordNet, Hu Liu, and SlangSD
sentiword <- hash_sentiment_sentiword
names(sentiword) <- c("word","score")

huliu <- hash_sentiment_huliu
names(huliu) <- c("word","score")

slangsd <- hash_sentiment_slangsd
names(slangsd) <- c("word","score")

# open The Ilad text file and tidy it up with tidytext
iliad <- tibble(line = 1:length(read_lines("TheIliad.txt")),
                    text = c(read_lines("TheIliad.txt"))) %>%
  unnest_tokens(word, text)

# create sentiments based on each of the lexicons
sentiword_sentiment <- iliad %>% 
  inner_join(sentiword, by = "word") %>%
  group_by(index = line %/% 80) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative"))

huliu_sentiment <- iliad %>% 
  inner_join(huliu, by = "word") %>%
  group_by(index = line %/% 80) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative"))

slangsd_sentiment <- iliad %>% 
  inner_join(slangsd, by = "word") %>%
  group_by(index = line %/% 80) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative"))

# plot each lexicon sentiment
ggplot(sentiword_sentiment, aes(index, sentiment, fill = overall)) +
  geom_col(show.legend = FALSE) + labs(title = "SentiWordNet")

ggplot(huliu_sentiment, aes(index, sentiment, fill = overall)) +
  geom_col(show.legend = FALSE) + labs(title = "Hu Liu")

ggplot(slangsd_sentiment, aes(index, sentiment, fill = overall)) +
  geom_col(show.legend = FALSE) + labs(title = "slangSD")
```

After looking at the sentiment of *The Iliad* using four different sentiment lexicons, I found it incredible how different the lexicons rated the same text. A text such as *The Iliad* should certainly come out as negative, but it was scired as neutral (*Hu Liu* and *SentiWordNet*) and negative (*slangSD*) depending on the sentiment lexicon used.  

I decided to remove stop words from the sentiments to see what impact those might be having.
```{r}
# remove stop words from each sentiment
sentiword_sentiment_go <- iliad %>% 
  inner_join(sentiword, by = "word") %>%
  anti_join(stop_words, by = "word") %>%
  group_by(index = line %/% 80) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative"))

huliu_sentiment_go <- iliad %>% 
  inner_join(huliu, by = "word") %>%
  anti_join(stop_words, by = "word") %>%
  group_by(index = line %/% 80) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative"))

slangsd_sentiment_go <- iliad %>% 
  inner_join(slangsd, by = "word") %>%
  anti_join(stop_words, by = "word") %>%
  group_by(index = line %/% 80) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative"))

# plot each lexicon sentiment
ggplot(sentiword_sentiment_go, aes(index, sentiment, fill = overall)) +
  geom_col(show.legend = FALSE) + labs(title = "SentiWordNet - no stop words")

ggplot(huliu_sentiment_go, aes(index, sentiment, fill = overall)) +
  geom_col(show.legend = FALSE) + labs(title = "Hu Liu - no stop words")

ggplot(slangsd_sentiment_go, aes(index, sentiment, fill = overall)) +
  geom_col(show.legend = FALSE) + labs(title = "slangSD - no stop words")
```

All of these plots made sense for a story such as *The Iliad*, but I was curious now about which words were the largest contributors to the overall sentiment of the story.
```{r}
iliad_words <- iliad %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  filter(n > 250)

ggplot(iliad_words, aes(x= reorder(word, n), y= n)) + geom_bar(stat = "identity", show.legend = FALSE) + coord_flip()

sentiword_words <- iliad %>% 
  inner_join(sentiword, by = "word") %>%
  anti_join(stop_words, by = "word") %>%
  group_by(word) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative")) %>%
  filter(abs(sentiment) > 50)

huliu_words <- iliad %>% 
  inner_join(huliu, by = "word") %>%
  group_by(word) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative")) %>%
  filter(abs(sentiment) > 150)

slangsd_words <- iliad %>% 
  inner_join(slangsd, by = "word") %>%
  anti_join(stop_words, by = "word") %>%
  group_by(word) %>% 
  summarise(sentiment = sum(score)) %>%
  mutate(overall = ifelse(sentiment > 0, "Positive", "Negative")) %>%
  filter(abs(sentiment) > 50)

ggplot(sentiword_words, aes(x= reorder(word, -abs(sentiment)), y= sentiment, fill = overall)) + geom_bar(stat = "identity", show.legend = FALSE) + labs(title = "SentiWordNet", x = "Word", y = "Impact")

ggplot(huliu_words, aes(x= reorder(word, -abs(sentiment)), y= sentiment, fill = overall)) + geom_bar(stat = "identity", show.legend = FALSE) + labs(title = "Hu Liu", x = "Word", y = "Impact")

ggplot(slangsd_words, aes(x= reorder(word, -abs(sentiment)), y= sentiment, fill = overall)) + geom_bar(stat = "identity", show.legend = FALSE) + labs(title = "slangSD", x = "Word", y = "Impact")

mean(sentiword_sentiment_go$sentiment) - mean(sentiword_sentiment$sentiment)
mean(huliu_sentiment_go$sentiment) - mean(huliu_sentiment$sentiment)
mean(slangsd_sentiment_go$sentiment) - mean(slangsd_sentiment$sentiment)
```

# Conclusion
After looking at the sentiment of *The Iliad* using three different sentiment lexicons, both with and without stop words, I find it incredible how different these lexicons are. Some lexicons (*slangSD*) swayed based on stop words, while others (*SentiWordNet*) had minimal change when stop words were removed.

After looking at the most impactful words from each sentiment lexicon, I believe the *Hu Liu* lexicon is the most approriate for this text. The most impactful words are meaningful even before stop words are removed and the overall sentiment outcome looks similar to what I would expect.

#### **Citations**
1. Silge, Julia, and David Robinson. “Text Mining with R.” *2 Sentiment Analysis with Tidy Data*, 7 Mar. 2020, www.tidytextmining.com/sentiment.html.  
2. Baccianella S., Esuli, A. and Sebastiani, F. (2010). SentiWordNet 3.0: An Enhanced Lexical Resource for Sentiment Analysis and Opinion Mining. International Conference on Language Resources and Evaluation.  
3. Hu, M., & Liu, B. (2004). Mining and summarizing customer reviews. Proceedings of the ACM
SIGKDD International Conference on Knowledge Discovery & Data Mining (KDD-2004). Seattle,
Washington.
Hu, M., & Liu, B. (2004). Mining opinion features in customer reviews. National Conference on
Artificial Intelligence.
4. Wu, L., Morstatter, F., and Liu, H. (2016). SlangSD: Building and using a sentiment dictionary of
slang words for short-text sentiment classification. CoRR. abs/1168.1058. 1-15.