---
title: "XML and JSON"
author: "David Moste"
date: "3/15/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
For this assignment, I created an html, xml, and json file containing information about three of my college physics textbooks. These files contained the book titles, authors, subject, and page counts. I then used R to read each file and create a dataframe from the information.

### Code
I read in each file using the XML, jsonlite, and rvest libraries. Then I converted them to separate dataframes.
```{r}
library(XML)
library(jsonlite)
library(rvest)

html <- as.data.frame(read_html("books.html") %>% html_table())

xml <- xmlParse(file = "books.xml")
xml <- xmlRoot(xml)
xml <- xmlToDataFrame(xml)

json <- fromJSON("books.json")
json <- json$books

html
xml
json
```

# Conclusion
Each of the dataframes is a little bit different than the others. The HTML dataframe lost the headers for the columns and the JSON dataframe represented the authors as a list instead of a single string (which is nice).
